---
title: 'Part 1: R Reference Guide'
author: "E. Baylor"
date: "12/16/2021"
output: html_document
---
# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages
```{r}
require(here)
library(palmerpenguins)

ginkgo <- data.frame(
  read.csv(here("data", "ginkgo_data_2021.csv"))
)
```


## Data Structures

### C()
The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

All of the elements must be of the same type.
I can’t combine character and numeric types in the same call to c()
Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

##Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

```
I can show the contents of a vector by typing the name of the vector, or using the print() function.
```{r}
## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)
```

### Length ()

The function length() is used to determine the ength of a vector of factor.
```{r}
length(ginkgo$petiole_length)
```

### Matrix()

The function Matrix() creates the Matrix from data values.
```{r}
matrix(data = 0:70, nrow = 7, ncol = 7, byrow = FALSE)
```

### Data.frame()

The function data.frame() is a list of two or more related variables.
```{r}
head(data.frame(ginkgo))

```
```{r}
nrow(ginkgo)
ncol(ginkgo)
dim(ginkgo)
```

## Subsetting

### $
```{r}
head(ginkgo$petiole_length)
```

### []
```{r}
ginkgo[1,]

ginkgo[2,3]

head(ginkgo[,3])
```

### Subset()
```{r}
head(subset(penguins, species == "Adelie"))
```

## Numerical Data Exploration

### Summary()
The function Summary() is used to get the summary of various model fitting functions.
```{r}
summary(ginkgo)
```

### Mean()
The function mean() is used to calculate the average of a list.
```{r}
mean(ginkgo$notch_depth)
```

### sd()
The function sd() is used to determine the standard deviation of any given values in x.
```{r}
sd(ginkgo$notch_depth)
```

## Graphical Data Exploration

### Scatterplot()
```{r}
plot(max_width ~ max_depth, data = ginkgo, 
     xlab = "Max Leaf Depth", ylab = "Max Lead Width",
     main = "Scatterplot of Ginkgo Leaf Max Depth x Max Width",
     col = "#CD8162", xlim = c(10, 150), ylim = c(10, 150), pch= 2, cex = .5)
```

### Histogram()
```{r}
require(palmerpenguins)
penguins <- data.frame(penguins)

hist(penguins$flipper_length_mm, breaks = 10, main = "Histogram of Penguin Flipper Lengths", xlab = "Flipper Length (mm)", col = "lightpink")
```

### Boxplot()
```{r}
boxplot(ginkgo$petiole_length, xlab = "Petiole Length", main = "Boxplot of Petiole Length", col = "#9BCD9B")
```
```{r}
boxplot(ginkgo$seeds_present, ginkgo$max_depth, names = c("No Seeds", "Seed Presence"), main = "Boxplot of Seed Presence by the Max Leaf Depth", col = "#009ACD")
```

### Par() and Mrow of Histograms
```{r}
par(mfrow = c(2, 2))
hist(ginkgo$max_depth, main = "Hist of Max Depth", xlab = "Max Depth", col = "#7CCD7C")

hist(ginkgo$notch_depth, main = "Hist of Notch Depth", xlab = "Notch Depth", col = "#CDC1C5")

hist(ginkgo$petiole_length, main = "Hist of Petiole Length", xlab = "Petiole Length", col = "#8B8386")

hist(ginkgo$max_width, main = "Hist of Max Width", xlab = "Max Width", col = "#548B54")

```

## Distribution Functions

### Dnorm()
The function dnorm() calculates the density of normal distribution.
```{r}
dnorm(x = -7:7, mean = 1, sd = 3) 
```

### Pnorm()
The function pnorm() calculates the distribution function of normal distribution.
```{r}
pnorm(0, mean = 0, sd = 3)
```

### Qnorm()
The function qnorm() calculates the quantile function of normal distribution.
```{r}
qnorm(0.10, mean = 0, sd = 3)
```

### Dbinom()
The function dbinom() calculates the density of binomial distribution.
```{r}
dbinom(x = 1:10, size = 10, prob = 0.3)
```

### Pbinom()
The function pbinom() calculates the distribution function of binomial distribution.
```{r}
pbinom(100, size = 10, prob = 0.3)
```

### Qbinom()
The function qbinom() calculates the quantile function of binomial distribution.
```{r}
qbinom(p = 0.77, size = 5, prob = 0.33)
```
